---
title: "Explore Vegetation Data"
author: "Whalen Dillon"
date: "June 26, 2015"
output: html_document
---

In this document I am exploring vegetation data from the Sonoma County plot network. These data were collected usin vegetation transects across each of the plots circa plot establishment (2003/2004) and in 2011. The DBH of alternate and non-host species were also recorded in 2006, 2012, and 2014.

Read in general vegetation data related to each plot
```{r read in data}
canopy_cover <- read.csv("analysis/data/canopy_cover_qry.csv")
veg_transects <- rbind(read.csv("analysis/data/veg_transects_subset_qry_pre2011.csv"), read.csv("analysis/data/veg_transects_2011.csv"))
alt_sp_dbh <- read.csv("analysis/data/alt_species_dbh.csv")
```

## Canopy Cover - Densiometers

```{r examine canopy cover data}
library(plyr); library(dplyr)
str(canopy_cover)
head(canopy_cover)
# I am interested in the date and total estimated cover fore each plot
canopy_cover <- canopy_cover %>% select(PlotID, Date, PctTotalCover)
canopy_cover$Date <- as.Date(as.character(canopy_cover$Date), format = "%m/%d/%Y")
canopy_cover <- rename(canopy_cover, plotid = PlotID, date = Date, pct_cvr = PctTotalCover)
summary(canopy_cover)
filter(canopy_cover, is.na(date))
filter(canopy_cover, plotid == "APGAR01") # appears to be a duplicate entry...
canopy_cover <- filter(canopy_cover, is.na(date) == "FALSE")

library(lubridate)
canopy_cover$year <- year(canopy_cover$date)
canopy_cover <- select(canopy_cover, -date)
canopy_cover$plotid <- tolower(canopy_cover$plotid)
write.csv(canopy_cover, "analysis/data/plot_canopy_cover_dens.csv")
```

The subset veg transect data `veg2` is approximately half of the original data points, selecting every other sample point along the transect. About twice as many points were sampled during plot establishment compared to the 2011 sampling
```{r examine veg transect data}
str(veg_transects)
veg_transects$Date <- as.Date(as.character(veg_transects$Date), format = "%m/%d/%Y")
veg_transects$year <- year(veg_transects$Date)
summary(veg_transects)
head(veg_transects, 20)
unique(veg_transects$year)

dim(filter(veg_transects, year != 2011))
unique(filter(veg_transects, year == 2005 | year == 2004 | year == 2003)$PlotID) # 205 plot IDs
unique(filter(veg_transects, year == 2011)$PlotID) # 199 plot IDs
veg_transects$PlotID <- tolower(veg_transects$PlotID)

detach("package:lubridate", unload = T)
```

## Overstory Vegetation

```{r overstory veg}
veg_os <- select(veg_transects, -UnderstorySp)
summary(veg_os)

# veg_us <- veg_us %>% (UnderstorySp != "")
# summary(veg_us)

# veg_os <- veg_os %>% filter(OverstorySp != "")
veg_os[veg_os == -9999.00] <- NA
filter(veg_os, is.na(OverstorySp))
#veg_os <- na.omit(veg_os)
summary(veg_os) # Percent coverage max is greater than 100
filter(veg_os, PercentCoverage > 100)
veg_os$PercentCoverage[veg_os$PercentCoverage > 100] <- 80
summary(veg_os)

veg_os <- as.tbl(veg_os)
(veg_os)
veg_os$OverstorySp <- tolower(veg_os$OverstorySp)
unique(veg_os$OverstorySp)
veg_os$OverstorySp[veg_os$OverstorySp == "arcto sp."] <- "arcto"
veg_os$OverstorySp[veg_os$OverstorySp == "quke x quag"] <- "quag x quke"
veg_os$OverstorySp[veg_os$OverstorySp == "none"] <- "open"
veg_os$OverstorySp[veg_os$OverstorySp == "visp"] <- "vinca sp."
veg_os$OverstorySp[veg_os$OverstorySp == ""] <- NA
filter(veg_os, OverstorySp == "uq1")
unique((filter(veg_os, PlotID == "sugar04", year == 2004))$OverstorySp)
unique((filter(veg_os, PlotID == "sugar04", year == 2011))$OverstorySp)
# This hints at the "uq1" from 2004 being the hybrid recorded in 2011. Note that the "open" value will be dropped later, but right now is being used as a place holder.
veg_os$OverstorySp[veg_os$OverstorySp == "uq1"] <- "quke x quwi"
# I'm also making the assumption of the "quwi" recorded in 2004 also actually being a hybrid because we didn't record a "quwi" in 2011
veg_os$OverstorySp[veg_os$PlotID == "sugar04" & veg_os$year == 2004 & veg_os$OverstorySp == "quwi"] <- "quke x quwi"
unique(veg_os$OverstorySp)
# What's the deal with "unknown sp." & "unknown sp. 1"?
filter(veg_os, OverstorySp == "unknown sp.")
filter(veg_os, OverstorySp == "unknown sp. 1")
# I think I can assume that the "unknowns" are the same within a plot, but probably different across the plot network. So, I can use them for calculating richness or diversity within a plot, but should probably exclude them from any other metric.

# I want to get this arranged so that I have the number of observations for each species in each plot

veg_os_summ <- left_join(veg_os %>% filter(OverstorySp != "OPEN") %>% 
      group_by(PlotID, year, OverstorySp) %>% 
      summarise(ptct = length(OverstorySp)),
      veg_os %>% group_by(PlotID, year) %>% summarise(totpts = length(unique(PtNum)) * length(unique(Transect))),
      by = c("PlotID", "year"))
summary(veg_os_summ)
```

Why is the minimum for `totpts` 45 and the maximum 80? All *should* be 60, though I would allow for a little wiggle room. If I can't figure anything out then I'm going to call it good enough
```{r overstory outliers}
filter(veg_os_summ, totpts == 45)
filter(veg_os_summ, PlotID == "ann25")
(unique(filter(veg_os, PlotID == "ann25", year == 2004)$Transect))
(unique(filter(veg_os, PlotID == "ann25", year == 2004)$PtNum))
(unique(filter(veg_transects, PlotID == "ann25", year == 2004)$PtNum))

filter(veg_os_summ, PlotID == "mutte02")
unique(filter(veg_os, PlotID == "mutte02", year == 2004)$Transect)
unique(filter(veg_os, PlotID == "mutte02", year == 2011)$Transect)
(unique(filter(veg_os, PlotID == "mutte02", year == 2004)$PtNum))
(unique(filter(veg_transects, PlotID == "mutte02", year == 2004)$PtNum))
(unique(filter(veg_os, PlotID == "mutte02", year == 2011)$PtNum))
(unique(filter(veg_transects, PlotID == "mutte02", year == 2011)$PtNum))

filter(veg_os_summ, PlotID == "sugar32")
unique(filter(veg_os, PlotID == "sugar32", year == 2004)$Transect)
unique(filter(veg_os, PlotID == "sugar32", year == 2004)$PtNum)
# I don't see much I can do about these.

filter(veg_os_summ, totpts == 80)
filter(veg_os_summ, PlotID == "jlsp05")
# The ratios from this sampling are nearly the same (within < 1%), so that should be fine.
```

Here I am adding a column that calculates the proportional coverage for each species based on the number of points they were observed in throughout the plot and the total number of points sampled in the plot.
```{r calculate overstory proportianal coverage at each point}
veg_os_summ <- mutate(veg_os_summ, pcov = ptct/totpts)
veg_os_summ
```

## Understory Vegetation

Now on to working through the understory veg data. I think this will actually be the more important because I will have DBH measurements for most of the species that would be observed in the overstory. Though there will as always be a few exceptions, e.g. TODI.
```{r understory veg}
veg_us <- select(veg_transects, -OverstorySp, -PercentCoverage)
summary(veg_us)

veg_us <- as.tbl(veg_us)
veg_us
veg_us$UnderstorySp <- tolower(veg_us$UnderstorySp)
summary(veg_us)
unique(veg_us$UnderstorySp)
# summary(filter(veg_us, UnderstorySp == "None"))
# summary(filter(veg_us, UnderstorySp == "OPEN"))
veg_us[veg_us == -9999] <- NA
veg_us$UnderstorySp[veg_us$UnderstorySp == "none"] <- "open"
veg_us$UnderstorySp[veg_us$UnderstorySp == "arcto sp."] <- "arcto"
filter(veg_us, UnderstorySp == "vasp")

unique(filter(veg_us, PlotID == "ann33", year == 2005)$UnderstorySp)
unique(filter(veg_us, PlotID == "ann33", year == 2011)$UnderstorySp)
filter(veg_us, UnderstorySp == "")
veg_us$UnderstorySp[veg_us$PlotID == "ann33" & veg_us$year == 2011 & veg_us$UnderstorySp == "losp"] <- "lohi"
veg_us$UnderstorySp[veg_us$PlotID == "ann33" & veg_us$year == 2011 & veg_us$UnderstorySp == "syal"] <- "symo"
veg_us$UnderstorySp[veg_us$PlotID == "ann33" & veg_us$year == 2011 & veg_us$UnderstorySp == "roca"] <- "rogy"

(unique(veg_us$UnderstorySp))

filter(veg_us, UnderstorySp == "bacspp") # Assuming this is a Baccharis species
veg_us$UnderstorySp[veg_us$UnderstorySp == "bacspp"] <- "bapi"

filter(veg_us, UnderstorySp == " to") # Maybe it's supposed to say "arcto" or "todi"?
veg_us$UnderstorySp[veg_us$UnderstorySp == " to"] <- "todi"
filter(veg_us, UnderstorySp == "4") # this is probably a typo; it should be "umca"
veg_us$UnderstorySp[veg_us$UnderstorySp == "4"] <- "umca"

veg_us$UnderstorySp[veg_us$UnderstorySp == "quke x quag"] <- "quag x quke"

filter(veg_us, UnderstorySp == "losp")
summary(filter(veg_us, UnderstorySp == "losp"))
unique(filter(veg_us, UnderstorySp == "losp")$PlotID)
unique(filter(veg_us, PlotID == "ann14", year == 2011)$UnderstorySp)
unique(filter(veg_us, PlotID == "ann14", year == 2003)$UnderstorySp)
# I wonder if all the "losp" we did in 2011 are reallly just "lohi"
unique(filter(veg_us, year < 2011)$UnderstorySp)
unique(filter(veg_us, year >= 2011)$UnderstorySp)
filter(veg_us, year < 2011, UnderstorySp == "losp")
filter(veg_us, year < 2011, UnderstorySp == "lonicera sp.")
# Changing all to lohi
veg_us$UnderstorySp[veg_us$UnderstorySp == "losp"] <- "lohi"

filter(veg_us, UnderstorySp == "quag x")
unique(filter(veg_us, PlotID == "mcneil01", year == 2005)$UnderstorySp)
unique(filter(veg_us, PlotID == "mcneil01", year == 2011)$UnderstorySp)
# McNeil01 was decommissioned prior to 2011, so not sure what the "quag x" was. But if this is in fact a quag-hybrid, then it there should probably be an adult tagged stem. Going to check the database...this plot only had adult UMCA tagged. So, I'm going to assume this was just a (perhaps funny looking) quag seedling/sapling.
veg_us$UnderstorySp[veg_us$UnderstorySp == "quag x"] <- "quag"

unique((veg_us$UnderstorySp))
filter(veg_us, UnderstorySp == "todye") # assume that this should be "todi"
veg_us$UnderstorySp[veg_us$UnderstorySp == "todye"] <- "todi"

filter(veg_us, UnderstorySp == "arsp?") # assume "arcto" b/c ARME would be pretty obvious for ID
veg_us$UnderstorySp[veg_us$UnderstorySp == "arsp?"] <- "arcto"

filter(veg_us, UnderstorySp == "unk1")
filter(veg_us, UnderstorySp == "unknown sp. 1")
veg_us$UnderstorySp[veg_us$UnderstorySp == "unk1"] <- "unknown sp. 1"
veg_us$UnderstorySp[veg_us$UnderstorySp == "unk2"] <- "unknown sp. 2"

filter(veg_us, UnderstorySp == "uq1")
filter(veg_us, UnderstorySp == "us#1")
veg_us$UnderstorySp[veg_us$UnderstorySp == "uq1" | veg_us$UnderstorySp == "us#1" | veg_us$UnderstorySp == "unknown sp."] <- "unknown sp. 1"

veg_us$UnderstorySp[veg_us$UnderstorySp == "vinca sp"] <- "vinca sp."
veg_us$UnderstorySp[veg_us$UnderstorySp == "lusp" | veg_us$UnderstorySp == "lupinus sp"] <- "lupine sp."
veg_us$UnderstorySp[veg_us$UnderstorySp == "vasp"] <- "vaccinium sp."
veg_us$UnderstorySp[veg_us$UnderstorySp == "gasp"] <- "garrya sp."
veg_us$UnderstorySp[veg_us$UnderstorySp == "qusp"] <- "quercus sp."
veg_us$UnderstorySp[veg_us$UnderstorySp == "risp" | veg_us$UnderstorySp == "ribes sp"] <- "ribes sp."
veg_us$UnderstorySp[veg_us$UnderstorySp == "prunus sp"] <- "prunus sp."
veg_us$UnderstorySp[veg_us$UnderstorySp == "rosa" | veg_us$UnderstorySp == "rosp"] <- "rose sp."
veg_us$UnderstorySp[veg_us$UnderstorySp == "rubu sp."] <- "rusp"
veg_us$UnderstorySp[veg_us$UnderstorySp == "rhsp"] <- "rhamnus sp."
veg_us$UnderstorySp[veg_us$UnderstorySp == "garry sp."] <- "garrya sp."

filter(veg_us, UnderstorySp == "asca") # I suspect this was supposed to be "aeca", even thou "asca" is written on the data sheet
filter(veg_us, PlotID == "bouv03", PtNum == 6)
unique(filter(veg_us, PlotID == "bouv03", year == 2005)$UnderstorySp)
unique(filter(veg_us, PlotID == "bouv03", year == 2011)$UnderstorySp)
veg_us$UnderstorySp[veg_us$UnderstorySp == "asca"] <- "aeca"
```

Okay, I think I have solved all (most?) of the duplicate species issues. Now to summarise the understory vegetation data similar to the overstory data.
```{r summarise veg understory data}
veg_us_summ <- left_join(veg_us %>% filter(UnderstorySp != "", UnderstorySp != "open") %>% 
      group_by(PlotID, year, UnderstorySp) %>% 
      summarise(ptct = length(UnderstorySp)),
      veg_us %>% group_by(PlotID, year) %>% 
            summarise(totpts = length(unique(PtNum)) * length(unique(Transect))),
      by = c("PlotID", "year"))
# filter(veg_us, PlotID == "ANN05", year == 2011)

veg_us_summ <- mutate(veg_us_summ, pcov = ptct/totpts)
veg_us_summ
summary(veg_us_summ)
unique(veg_us_summ$UnderstorySp)

veg_us_2011 <- filter(veg_us_summ, year == 2011)
veg_us_2005 <- filter(veg_us_summ, year < 2011)

write.csv(veg_us_2005, "analysis/data/understory_veg_2005.csv")
write.csv(veg_us_2011, "analysis/data/understory_veg_2011.csv")
```

# Untagged Overstory
Now I need to examine the data for alternate and non-host species DBH. These are the stems that are untagged and meet a 5cm DBH threshold.

```{r untagged species}
summary(alt_sp_dbh)
library(lubridate)
alt_sp_dbh$date <- as.Date(alt_sp_dbh$Date, format = "%m/%d/%Y")
filter(alt_sp_dbh, is.na(date)) # this should be 5/22/2012
alt_sp_dbh$Date[alt_sp_dbh$Date == ""] <- "5/22/2012 0:00:00"
alt_sp_dbh$date <- as.Date(alt_sp_dbh$Date, format = "%m/%d/%Y")
summary(alt_sp_dbh)
alt_sp_dbh$year <- year(alt_sp_dbh$date)
head(alt_sp_dbh)

alt_sp_dbh <- alt_sp_dbh %>% select(-PlotVisit_PlotVisitID, -Date) %>% 
      rename(dbh = Measure)
unique(as.character(alt_sp_dbh$Species))

alt_sp_dbh$Species[alt_sp_dbh$Species == "PSME"] <- "psme"
alt_sp_dbh$Species <- tolower(alt_sp_dbh$Species)
summary(alt_sp_dbh)
unique(alt_sp_dbh$Species)
unique(alt_sp_dbh$Dead)
unique(alt_sp_dbh$year)
# alt_sp_dbh$Dead <- ifelse(alt_sp_dbh$Dead == 0, "no", "yes")
alt_sp_dbh <- as.tbl(alt_sp_dbh)

summary(filter(alt_sp_dbh, year == 2003))
unique(droplevels(filter(alt_sp_dbh, year == 2003)$PlotID))# 32 plots
pid2003 <- as.character(unique(filter(alt_sp_dbh, year == 2003)$PlotID))
alt_sp_2003 <- filter(alt_sp_dbh, year == 2003)

summary(filter(alt_sp_dbh, year == 2004))
unique(droplevels(filter(alt_sp_dbh, year == 2004)$PlotID))# 120 plots
pid2004 <- as.character(unique(filter(alt_sp_dbh, year == 2004)$PlotID))
alt_sp_2004 <- filter(alt_sp_dbh, year == 2004)

pid2003 %in% pid2004
pid2004[47]; pid2004[69]
filter(alt_sp_dbh, year == 2003, PlotID == "MITSU02")
filter(alt_sp_dbh, year == 2004, PlotID == "MITSU02")

summary(filter(alt_sp_dbh, year == 2005))
unique(droplevels(filter(alt_sp_dbh, year == 2005)$PlotID))# 25 plots
pid2005 <- as.character(unique(filter(alt_sp_dbh, year == 2005)$PlotID))
alt_sp_2005 <- filter(alt_sp_dbh, year == 2005)

# I think I should work this backwards so that the most recent sampling is added to
pid2005 %in% pid2004# COOK02 is the only duplicate in the 2004 sampling
filter(alt_sp_2004, PlotID=="COOK02")
filter(alt_sp_2005, PlotID=="COOK02")
# But there are a lot more records in the 2004 data, which is closer to the 2012 & 2014 sampling. I could combine the 2004 & 2005 data, including the potentially duplicated COOK02 data.
# summary(filter(alt_sp_2012, PlotID=="COOK02"))
# filter(alt_sp_2014, PlotID=="COOK02")

alt_sp_45 <- rbind(alt_sp_2005, alt_sp_2004)
summary(alt_sp_45)
pid45 <- as.character(unique(alt_sp_45$PlotID))# 144 unique plots
pid45 %in% pid2003
pid45[71]; pid45[93]; pid45[133]
filter(alt_sp_45, PlotID == "MITSU02")
filter(alt_sp_2003, PlotID == "MITSU02")
# Exclude MITSU02 from the 2003 data
alt_sp_2003 <- filter(alt_sp_2003, PlotID != "MITSU02")

filter(alt_sp_45, PlotID == "LINS01")
filter(alt_sp_2003, PlotID == "LINS01")
# Keep LINS01 from 2003 data, add to other data
filter(alt_sp_45, PlotID == "GANAY02")
filter(alt_sp_2003, PlotID == "GANAY02")
# Exclude GANAY02 from the 2004, keep the 2003 data
alt_sp_45 <- filter(alt_sp_45, PlotID != "GANAY02")
unique(alt_sp_2003$PlotID) %in% unique(alt_sp_45$PlotID)# LINS01 still there
alt_sp_45 <- rbind(alt_sp_45, alt_sp_2003)
alt_sp_2005 <- alt_sp_45
rm(alt_sp_45)

summary(filter(alt_sp_dbh, year == 2012))
pid2012 <- unique(droplevels(filter(alt_sp_dbh, year == 2012)$PlotID))# 171 plots
alt_sp_2012 <- filter(alt_sp_dbh, year == 2012)

summary(filter(alt_sp_dbh, year == 2014))
pid2014 <- unique(droplevels(filter(alt_sp_dbh, year == 2014)$PlotID))# 164 plots
alt_sp_2014 <- filter(alt_sp_dbh, year == 2014)

# Determine which plots from 2012 (the one with more records) aren't in the 2014 data.
pid2012 %in% pid2014
pid2012sub <- droplevels(pid2012[c(4,12:13,20:21,52:53,111,114)])
alt_sp_2012 <- filter(alt_sp_2012, PlotID == "ARBIT04" | PlotID == "LINS01" | PlotID == "LUPIN01" | PlotID == "MITSU02" | PlotID == "MITSU04" | PlotID == "SKILE01" | PlotID == "SPAUL01" | PlotID == "GANAY02" | PlotID == "HALOW02")

alt_sp_2014 <- rbind(alt_sp_2014, alt_sp_2012)
summary(alt_sp_2014)
alt_sp_2014$year <- 2014
summary(alt_sp_2005)
alt_sp_2005$year <- 2005

alt_sp_2014 <- select(alt_sp_2014, -date)
alt_sp_2005 <- select(alt_sp_2005, -date)

alt_species <- rbind(alt_sp_2005, alt_sp_2014)
sort(unique(alt_species$species))
alt_species[alt_species=="arma"] <- "arcto"
write.csv(alt_species, "analysis/data/alt-species-dbh-corrected.csv", row.names = F)
```

```{r group alt species dbh to plot and year}
alt_species
alt_species$Dead[alt_species$Dead==0] <- "Alive"
alt_species$Dead[alt_species$Dead==1] <- "Dead"
alt_species <- rename(alt_species, status = Dead, plotid = PlotID, species = Species)
alt_dbh_plot <- alt_species %>% 
      group_by(plotid, species, status, year) %>% 
      summarise(tot_dbh = sum(dbh, na.rm = T), avg_dbh = mean(dbh, na.rm = T),
                tot_ba_m2 = sum(pi*dbh^2/40000, na.rm = T),
                avg_ba_m2 = mean(pi*dbh^2/40000, na.rm = T),
                abundance = length(dbh))
write.csv(alt_dbh_plot, "analysis/data/alt-dbh-plot.csv", row.names = F)
```


```{r }
# Create data frames of live and dead stems for '2005' and '2014' samplings. In reality the 2005 grouping has data from visits during 2003 and 2004, and the 2014 grouping includes data from some visits in 2012.
alt_sp_2014_live <- filter(alt_sp_2014, Dead == 0)
alt_sp_2014_live <- select(alt_sp_2014_live, -Dead)
alt_sp_2005_live <- filter(alt_sp_2005, Dead == 0)
alt_sp_2005_live <- select(alt_sp_2005_live, -Dead)
alt_sp_2014_dead <- filter(alt_sp_2014, Dead == 1)
alt_sp_2014_dead <- select(alt_sp_2014_dead, -Dead)
alt_sp_2005_dead <- filter(alt_sp_2005, Dead == 1)# only 12 obs, all at GANAY02
alt_sp_2005_dead <- select(alt_sp_2005_dead, -Dead)

# Confirm year assignments are correct/consistent.
unique(alt_sp_2014_live$year)
unique(alt_sp_2005_live$year)
unique(alt_sp_2005_dead$year)
unique(alt_sp_2014_dead$year)


alt_sp_summ_2005 <- left_join(
      # First data frame
      alt_sp_2005_live %>%
            group_by(PlotID, Species, year) %>% 
            summarise(
                  live_count = length(Species),
                  live_avg_dbh = mean(dbh),
                  live_tot_dbh = sum(dbh)),
      # Second data frame
      alt_sp_2005_dead %>%
            group_by(PlotID, Species, year) %>% 
            summarise(
                  dead_count = length(Species), 
                  dead_avg_dbh = mean(dbh), 
                  dead_tot_dbh = sum(dbh))
)
summary(alt_sp_summ_2005)

alt_sp_summ_2014 <- left_join(
      # First data frame
      alt_sp_2014_live %>%
            group_by(PlotID, Species, year) %>% 
            summarise(
                  live_count = length(Species),
                  live_avg_dbh = mean(dbh),
                  live_tot_dbh = sum(dbh)),
      # Second data frame
      alt_sp_2014_dead %>%
            group_by(PlotID, Species, year) %>% 
            summarise(
                  dead_count = length(Species), 
                  dead_avg_dbh = mean(dbh), 
                  dead_tot_dbh = sum(dbh))
)
summary(alt_sp_summ_2014)

# Write out summary data frames of the untagged DBH species
write.csv(alt_sp_summ_2005, "analysis/data/untagged_dbh2005_summary.csv", row.names = F)
write.csv(alt_sp_summ_2014, "analysis/data/untagged_dbh2014_summary.csv", row.names = F)
```

```{r make wide format species abundance}
library(tidyr)
alt_dbh_plot <- read.csv("analysis/data/alt-dbh-plot.csv")
# alt_species$PlotID <- tolower(alt_species$PlotID)
# unique(alt_species$species)
# alt_species$Species[alt_species$Species=="arcto sp."] <- "arcto"
# alt_species$Species[alt_species$Species=="prunus sp."] <- "prunus"
# alt_species$Species[alt_species$Species=="unknown sp. 1"] <- "unk1"
# alt_species$Species[alt_species$Species=="quercus sp."] <- "qusp"
# head(alt_species)

head(alt_dbh_plot)
unique(alt_dbh_plot$species)
alt_dbh_plot.wide <- as.tbl(alt_dbh_plot %>% 
      select(plotid, species, status, year, abundance) %>% 
      spread(species, abundance))

# unique(alt_sp_2014_live$Species)
# alt_sp_2014_live$Species[alt_sp_2014_live$Species=="arcto sp."] <- "arcto"
# alt_sp_2014_live$Species[alt_sp_2014_live$Species=="prunus sp."] <- "prunus"
# alt_sp_2014_live$Species[alt_sp_2014_live$Species=="unknown sp. 1"] <- "unk1"
# alt_sp_2014_live$Species[alt_sp_2014_live$Species=="quercus sp."] <- "qusp"
# 
# unique(alt_sp_2005_live$Species)
# alt_sp_2005_live$Species[alt_sp_2005_live$Species=="quercus sp."] <- "qusp"
# alt_sp_2005_live$Species[alt_sp_2005_live$Species=="arcto sp."] <- "arcto"
# alt_sp_2005_live$Species[alt_sp_2005_live$Species=="prunus sp."] <- "prunus"

# Aggregate to species counts by plot
alt_sp_05_live_abund <- alt_sp_2005_live %>% select(-dbh, -year) %>% 
      group_by(PlotID, Species) %>% summarise(count = length(Species))
# Spread from long to wide format
alt_sp_05_live_abund <- alt_sp_05_live_abund %>% spread(Species, count)

alt_sp_14_live_abund <- alt_sp_2014_live %>% select(-dbh, -year) %>% 
      group_by(PlotID, Species) %>% summarise(count = length(Species))
alt_sp_14_live_abund <- alt_sp_14_live_abund %>% spread(Species, count)

# Create species abundance data frames of dead species
alt_sp_05_dead_abund <- alt_sp_2005_dead %>% select(-dbh, -year) %>% 
      group_by(PlotID, Species) %>% summarise(count = length(Species))
alt_sp_05_dead_abund <- alt_sp_05_dead_abund %>% spread(Species, count)

alt_sp_14_dead_abund <- alt_sp_2014_dead %>% select(-dbh, -year) %>% 
      group_by(PlotID, Species) %>% summarise(count = length(Species))
alt_sp_14_dead_abund <- alt_sp_14_dead_abund %>% spread(Species, count)

# Write out wide format abundance data frames
write.csv(alt_sp_05_live_abund, "analysis/data/altsp05abund_live.csv")
write.csv(alt_sp_14_live_abund, "analysis/data/altsp14abund_live.csv")
write.csv(alt_sp_05_dead_abund, "analysis/data/altsp05abund_dead.csv")
write.csv(alt_sp_14_dead_abund, "analysis/data/altsp14abund_dead.csv")
```

