---
title: "IALE 2015 Analysis"
author: "Whalen Dillon"
date: "June 15, 2015"
output: html_document
---


```{r load data}
# Ecological data: leaf counts, infection
load("~/GitHub/superspreaders/stems_plots.RData")
# Vegetation data
veg_us_2005 <- read.csv("analysis/data/understory_veg_2005.csv")
veg_us_2011 <- read.csv("analysis/data/understory_veg_2011.csv")
# Untagged dbh data
untagged_dbh <- read.csv("analysis/data/untagged_dbh_summary.csv")

# Physical environment: topography
library(rgdal)
plots_shp <- readOGR("analysis/data/plot_gis/", "soco_plots_metric")
summary(plots_shp)
plot(plots_shp, axes=T)
plots_env <- plots_shp@data

# Temperature: rainy season hour summaries
library(plyr); library(dplyr)
plots_temps <- read.csv("analysis/data/rainy_season_temps.csv")
summary(plots_temps)
plots_temps <- select(plots_temps, -X)
```

Join the plot level environmental data (geographic/topographic & temperatures)
```{r join plot data, message=FALSE}
plots_env <- left_join(plots_env, plots_temps, by = c("PLOT_ID"="plotid"))
summary(plots_env)
str(plots_env)
plots_env <- plots_env %>% select(-cat) %>% rename(easting = X, northing = Y, plotid = PLOT_ID)
```

Investigating `NAs` in the `plots_env` data set
```{r investigate NAs}
summary(plots_env)
filter(plots_env, is.na(sample_year))

# BUSH01 plot included. I think we have limited ecological data for this plot, and no microclimate data. So, I am going to exclude it from all the data sets
plots_env <- filter(plots_env, plotid != "BUSH01")
stems <- rename(stems, plotid = plot) %>% filter(plotid != "BUSH01")

summary(plots_env) # A number of observations for above 25 threshold have NAs
# I believe that these are in fact truly zero observations. NAs occurred for the 2010, 2011, and 2014 seasons, and the only NAs in the data set are associated with the above 25 threshold. 
plots_env %>% filter(is.na(hrs_abv25))
# I am replacing the NAs with `0`, assuming true observations.
plots_env[is.na(plots_env)] <- 0
```

Examine the bay laurel data at individual and plot levels
```{r examine umca data}
summary(plots_umca)
# This only has number of infected, uninfected, and total, which must apply to living stems only. I want to analyze leaf counts.
summary(bay_laurel)
# This has stem-level data on bay laurel.
```

I can use this to calculate plot level leaf count variables (sum, average) for each plot and year
```{r calculate plot level leaf counts}
plots_umca <- left_join(plots_umca, bay_laurel %>% filter(status == "Alive") %>% 
      group_by(plot, year) %>% 
      summarise(
            avg_lfct = mean(slc, na.rm = T),
            tot_lfct = sum(slc, na.rm = T)),
      by = c("plot", "year"))
summary(plots_umca)
```

```{r examine lfct NAs}
filter(plots_umca, is.na(avg_lfct))
# I'm not sure why this wasn't excluded from the calculations with the na.rm = T option
filter(bay_laurel, plot == "YAHNG02", status == "Alive")
# I'm removing it manually from the `plots_umca` data frame
plots_umca <- plots_umca %>% filter(avg_lfct != "NaN")
summary(plots_umca)
```

So now I have a plot-level of UMCA infection levels for each year. I want to join this with some of the environmental data.

```{r join env and umca data}
summary(plots_env)
plots_env$avg_psi <- rowMeans(cbind(plots_env$psiSpr, plots_env$psiSum, plots_env$psiWin))
summary(plots_umca)
# plot == plotid, year == sample_year
plots_umca <- left_join(select(plots_umca, -species), plots_env, by = c("plot" = "plotid", "year" = "sample_year"))
summary(plots_umca)
filter(plots_umca, is.na(easting))
# NAs are due to no microclimate data for year 2003, and abandoned plots BUSH01 & SUGAR02
```

Now I also want to calculate some plot-level community variables from the vegetation data. First, understory richness from the vegetation transects.
```{r calculate understory richness}
summary(veg_us_2005)
veg_us_2005 <- as.tbl(select(veg_us_2005, -X))
summary(veg_us_2011)
veg_us_2011 <- as.tbl(select(veg_us_2011, -X))
veg_us_2011
veg_us_2005

veg <- rbind(veg_us_2005, veg_us_2011)

unique(veg$year)
veg$year[veg$year < 2011] <- 2005 # Creates the two most complete veg sampling
unique(veg$us_species)
veg <- veg %>% rename(us_species = UnderstorySp, plotid = PlotID)

# I want to match the species to those used by Sarah in her second chapter
local_woody <- read.csv("analysis/data/presence_local_woody_n197.csv")
summary(local_woody)
unique(local_woody$spp.local)
woody_spp <- as.character(local_woody$spp.local)
woody_spp[woody_spp == "prunus"] <- "prunus sp."
sort(woody_spp)
sort(unique(veg$us_species))
# It looks like I will want to add a few more to her list

filter(veg, us_species == "ribes sp.")
filter(veg, us_species == "ribes")
veg$us_species[veg$us_species == "ribes"] <- "ribes sp."

woody_spp <- c(woody_spp, "ceanothus sp.", "cebe", "cecu", "frla", "garrya sp.", "gemo", "juca", "mafu", "mane", "mimulus sp", "pisa", "quag x quke", "qube", "quercus sp.", "quwi", "rhamnus sp.", "rhcr", "ribes sp.", "rica", "risa", "roca", "rosa sp.", "rubus sp.", "rudi", "salix sp.", "syal", "vaccinium sp.")

veg_sub <- filter(veg, veg$us_species %in% woody_spp)

richness <- veg_sub %>%
      group_by(plotid, year) %>% 
      summarise(us_rich = length(unique(us_species)))
richness
```

Next, overstory richness from the DBH data for tagged & untagged stems.

```{r calculate overstory richness}
summary(untagged_dbh)
untagged_dbh <- as.tbl(untagged_dbh %>% select(-X) %>% rename(plotid = PlotID, species = Species))
untagged_dbh
unique(untagged_dbh$year)
untagged_dbh$year[untagged_dbh$year < 2014] <- 2005 # Create two most complete groups
untagged_dbh$plotid <- tolower(untagged_dbh$plotid)

untagged_dbh %>% group_by(plotid, year) %>% summarise(untagged_rich = length(unique(species)))

stems <- as.tbl(stems)
summary(stems)
unique(stems$status)

stems <- stems %>% select(-X) %>% filter(status == "Alive" | status == "Dead")
unique(stems$species)
stems$species <- tolower(stems$species)
stems$species[stems$species == "quke x quag"] <- "quag x quke"
stems$plotid <- tolower(stems$plotid)
stems %>% group_by(plotid, year) %>% summarise(tagged_rich = length(unique(species)))

os_richness <- left_join(
      stems %>% group_by(plotid, year) %>% summarise(tagged_rich = length(unique(species))),
      untagged_dbh %>% group_by(plotid, year) %>% summarise(untagged_rich = length(unique(species))), by = c("plotid", "year")
) %>% mutate(os_rich = tagged_rich + untagged_rich)
os_richness

richness <- left_join(os_richness, richness, by = c("plotid","year"))
richness
summary(richness)
```

Now I want to join the richness data set to the `plots_umca` data frame that currently has the environmental and leaf count/infection variables. 

```{r join richness data}
plots_umca
richness
plots_umca <- rename(plots_umca, plotid = plot)
plots_umca$plotid <- tolower(plots_umca$plotid)

plots_umca <- left_join(plots_umca, richness, by = c("plotid", "year"))
```

Make a pairs plot of the continuous variables for UMCA plots.
```{r umca sod pairs plot}
# Helper functions for `pairs`
panel.hist <- function(x, ...)
{
      usr <- par("usr"); on.exit(par(usr))
      par(usr = c(usr[1:2], 0, 1.5) )
      h <- hist(x, plot = FALSE)
      breaks <- h$breaks; nB <- length(breaks)
      y <- h$counts; y <- y/max(y)
      rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
      usr <- par("usr"); on.exit(par(usr))
      par(usr = c(0, 1, 0, 1))
      r <- abs(cor(x, y, use = "na.or.complete"))
      txt <- format(c(r, 0.123456789), digits = digits)[1]
      txt <- paste0(prefix, txt)
      if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
      text(0.5, 0.5, txt, cex = cex.cor * r)
}

plots_umca <- ungroup(plots_umca)

pairs(select(plots_umca, -plotid, -year, -starts_with("psi"), -tagged_rich, -untagged_rich), lower.panel = panel.cor, diag.panel = panel.hist, main = "UMCA SOD Data")
```

Now that I have summarized some vegetation (not species specific), environmental, and bay laurel disease data to the plot-level I want to join this to the plot-level oak infection data.

```{r join umca and qusp data}
plots_umca
plot_qusp # right now this is broken down by oak species
summary(plot_qusp)
oak_sod <- plot_qusp %>% group_by(plot, year) %>%
      summarise(uninf_oak_ct = sum(uninfected_oak_ct), inf_oak_ct = sum(infected_oak_ct), tot_oak_ct = sum(c(uninfected_oak_ct, infected_oak_ct)))
oak_sod
oak_sod <- rename(oak_sod, plotid = plot)
oak_sod$plotid <- tolower(oak_sod$plotid)

oak_sod <- left_join(oak_sod, plots_umca[1:7], 
                     by = c("plotid", "year"))
summary(oak_sod) # the NA's for the bay laurel count variables are legitimately zeroes b/c there are no bay in some of these plots
oak_sod$uninfected_bay_ct[is.na(oak_sod$uninfected_bay_ct)] <- 0
oak_sod$infected_bay_ct[is.na(oak_sod$infected_bay_ct)] <- 0
oak_sod$tot_bay[is.na(oak_sod$tot_bay)] <- 0

head(plots_env)
plots_env$plotid <- tolower(plots_env$plotid)
oak_sod <- left_join(oak_sod, plots_env, by = c("plotid", "year" = "sample_year"))
oak_sod

richness
oak_sod <- left_join(oak_sod, richness, by = c("plotid", "year"))
summary(oak_sod)
summary(filter(oak_sod, is.na(avg_tmax)))
filter(oak_sod, is.na(avg_tmax), year == 2005) # BUSH01 plot; I'm dropping that from the data set
oak_sod <- filter(oak_sod, plotid != "bush01")
summary(filter(oak_sod, is.na(avg_tmax)))
filter(oak_sod, is.na(avg_tmax), year == 2004) # PONTI01 and SWEET01 were other early abandoned plots
oak_sod <- filter(oak_sod, plotid != "ponti01", plotid != "sweet01")
summary(filter(oak_sod, is.na(avg_tmax))) # Now the NAs in the environmental variables are just for year 2003 when we had severely incomplete temperature data
summary(oak_sod)

filter(oak_sod, plotid == "ann05") # There is no 2007 record for ann05 in the UMCA form, but there is a plot visit. I will need to come back and investigate this further for correcting the database.
oak_sod$infected_bay_ct[oak_sod$plotid == "ann05" & oak_sod$year == 2007] <- 15
oak_sod$tot_bay[oak_sod$plotid == "ann05" & oak_sod$year == 2007] <- 15
# I filled in the counts based on the values in the immediately preceding and following years, which I think is pretty solid reasoning. However, there is nothing I can do about the leaf counts right now.
summary(oak_sod)

filter(oak_sod, year == 2003) # 87 records
filter(oak_sod, year == 2004) # 90 records
# 2005 appears to be the first full year of observing all the plots
filter(oak_sod, year == 2005) # 171 records
filter(oak_sod, year == 2006) # 172 records
filter(oak_sod, year == 2007) # 171 records
filter(oak_sod, year == 2008) # 171 records
filter(oak_sod, year == 2009) # 170 records
filter(oak_sod, year == 2010) # 170 records
filter(oak_sod, year == 2011) # 170 records; final year of Sarah's 2nd chapter
filter(oak_sod, year == 2012) # 169 records
filter(oak_sod, year == 2014) # 168 records
# In Sarah's analysis we end up using only 163 plots
```

Now to do a pairs plot of the `oak_sod` data frame

```{r oak_sod pairs plot}
class(oak_sod)
oak_sod <- ungroup(oak_sod)
pairs(select(oak_sod, -plotid, -year, -starts_with("psi"), -tagged_rich, -untagged_rich), lower.panel = panel.cor, diag.panel = panel.hist, main = "Oak_SOD Data")
```



I also want to calculate some variables related to DBH, and in particular the "importance value" for each species. Importance value is calculated in the literature as:

`IV = 0.5 * (relative density / relative dominance)` where 
`Relative Density = # stems of species / total # of stems` and
`Relative Dominance = basal area of species / total basal area`

```{r calculate importance value}
unique(untagged_dbh$species)
untagged_dbh
unique(stems$species)

tagged_dbh <- left_join(stems %>% 
      filter(status == "Alive", location == "In") %>%
      select(plotid, year, species, dbh) %>% 
      group_by(plotid, species, year) %>% 
      summarise(live_count = length(species), live_avg_dbh = mean(dbh), live_tot_dbh = sum(dbh)),
      stems %>% 
      filter(status == "Dead", location == "In") %>%
      select(plotid, year, species, dbh) %>% 
      group_by(plotid, species, year) %>% 
      summarise(dead_count = length(species), dead_avg_dbh = mean(dbh), dead_tot_dbh = sum(dbh)),
      by = c("plotid", "species", "year")
)
tagged_dbh <- as.tbl(tagged_dbh)
summary(tagged_dbh)

tagged_dbh
untagged_dbh

stems_dbh <- rbind(tagged_dbh, untagged_dbh)

stems_dbh <- stems_dbh %>% group_by(plotid, year) %>% mutate(all_live_count = sum(live_count), all_live_dbh = sum(live_tot_dbh, na.rm = T))
summary(stems_dbh)
filter(stems_dbh, all_live_dbh > 800)

stems_dbh <- stems_dbh %>% mutate(rel_dens = live_count / all_live_count, rel_dom = live_tot_dbh / all_live_dbh, imp_value = 0.5*(rel_dens + rel_dom))
summary(stems_dbh)
write.csv(stems_dbh, "analysis/data/species_dbh_impval_plot.csv")
```